%%%% Reference file for:
%%%% Eduardo D. Sontag,
%%%% __Mathematical Control Theory: Deterministic Finite Dimensional Systems__
%%%% Springer, NY, 1990. (396+xiii pages, ISBN 0-387-97366-4; 3-540-97366-4)
%%%% Series: Textbooks in Applied Mathematics, Number 6.
%%%% Hardcover, $39.00
%%%% Order in USA from 1-800-SPRINGER (in NJ, 201-348-4033)
%%%% save into ``book.tex'', then compile as ``latex book.tex''

\documentstyle[11pt]{article}
\topmargin -0.5in
\textheight 8.8in
\oddsidemargin 0.0in
\evensidemargin 0.0in
\textwidth 6.7in

\begin{document}

\begin{center}
{\bf List of References for}\\
{\it Mathematical Control Theory: Deterministic Finite Dimensional Systems}\\
by Eduardo D. Sontag, \copyright Springer Verlag 1990, NY\\
Series: Textbooks in Applied Mathematics, Number 6
\end{center}
\vspace{1.0cm}
\begin{enumerate}

\item
Abed, E.H., and J-H.\ Fu, ``Local stabilization and bifurcation control, I.
Hopf bifurcation,'' {\it Systems and Control Letters} {\bf 7}(1986): 11-17.

\item
Abraham, R., and J.E.\ Marsden, {\it Foundations of Mechanics, second
edition}, Benjamin Cummings, Reading, 1978.

\item
Adamjan, V.M., D.Z.\ Arov, and M.G.\ Krein, ``Analytic properties of Schmidt
pairs for a Hankel operator and the generalized Schur-Takagi problem,'' {\it
Math.\ USSR Sbornik} {\bf 15}(1971): 31-73.

\item
Aeyels, D., ``Stabilization of a class of nonlinear systems by a smooth
feedback control,'' {\it Systems and Control Letters} {\bf 5}(1985): 289-294.

\item
Aeyels, D., ``Generic observability of differentiable systems,'' {\it SIAM
J.Control \& Opt.} {\bf 19}(1981): 595-603.

\item
Aeyels, D., ``On the number of samples necessary to achieve observability,''
{\it Systems and Control Letters} {\bf 1}(1981): 92-94.

\item
Aeyels, D., ``Global observability of Morse-Smale vector fields,'' {\it J.\
Diff.\ Eqs.} {\bf 45}(1982): 1-15.

\item
Ahmed, N.U., {\it Elements of Finite-Dimensional Systems and Control Theory},
Wiley, New York, 1988.

\item
Akaike, H., ``Stochastic theory of minimal realization,'' {\it IEEE Trans.\
Autom.\ Control} {\bf AC-19}(1974): 667-674.

\item
Akaike, H., ``Markovian representation of stochastic processes by canonical
variables,'' {\it SIAM J.\ Control \& Opt.}{\bf 13}(1975):162-173.

\item
Anderson, B.D.O., and J.B.\ Moore, {\it Linear Optimal Control},
Prentice-Hall, Englewood Cliffs, 1971.

\item
Antoulas, A.C., ``On recursiveness and related topics in linear systems,''
{\it IEEE Trans.\ Autom.\ Control} {\bf AC-31}(1986): 1121-1135.

\item
Antoulas, A.C., and B.D.O.\ Anderson, ``On the scalar rational interpolation
problem,'' {\it IMA J.\ Math.\ Control Infor.}{\bf 3}(1986):61-\-88.

\item
Aoki, M., {\it Optimal Control and System Theory in Dynamic Economic
Analysis}, North Holland, New York, 1976.

\item
Arbib, M.A., ``A characterization of multilinear systems,'' {\it IEEE Trans.\
Autom.\ Control} {\bf 14}(1969): 699-702.

\item
Arbib, M.A., and E.\ Manes, ``Machines in a category: an expository
introduction,'' {\it SIAM Review} {\bf 57}(1974): 163-192.

\item
Artstein, Z., ``Stabilization with relaxed controls,'' {\it Nonl.\ Anal., TMA}
{\bf 7}(1983): 1163-1173.

\item
Astrom, K.J., {\it Introduction to Stochastic Control Theory}, Academic Press,
New York, 1970.

\item
Athans, M., and P.L.\ Falb, {\it Optimal Control: An Introduction to the
Theory and its Applications}, McGraw-Hill, New York, 1966.

\item
Baccioti, A., and P.\ Boieri, ``Linear stabilizability of planar nonlinear
systems,'' {\it Math.\ Control Signals Systems} {\bf 3}(1990): 183-193.

\item
Baillieul, J., ``Controllability and observability of polynomial dynamical
systems,'' {\it Nonl.Anal.,TMA} {\bf 5}(1981): 543-552.

\item
Baillieul, J., R.W.\ Brockett, and R.B.\ Washburn, ``Chaotic motion in
nonlinear feedback systems,'' {\it IEEE Trans.\ Circuits and Systems} {\bf
CAS-27}(1980): 990-997.

\item
Banks, S.P., {\it Control Systems Engineering : Modelling and Simulation,
Control Theory, and Microprocessor Implementation}, Prentice-Hall, Englewood
Cliffs, 1986.

\item
Banks, S.P., ``Stabilizability of finite- and infinite-dimensional bilinear
systems, {\it IMA J.\ of Math.\ Control and Inform.} {\bf 3}(1986): 255-271.

\item
Barmish, B.R., M.J.\ Corless, and G.\ Leitmann, ``A new class of stabilizing
controllers for uncertain dynamical systems,'' {\it SIAM J.  Control \& Opt.}
{\bf 21}(1983): 246-255.

\item
Barnett, S., {\it Polynomials and Linear Control Systems}, Marcel Dekker, New
York, 1983.

\item
Barnett, S., {\it Matrices in Control Theory}, Krieger, Malabar, Fla., 1984.

\item
Barnett, S., {\it Introduction to Mathematical Control Theory}, Oxford
University Press, New York, 1985.

\item
Bartlett, A.C., C.V.\ Hollot, and H.\ Lin, ``Root locations of an entire
polytope of polynomials: It suffices to check the edges,'' {\it Math.\ Control
Signals Systems} {\bf 1}(1988): 61-71.

\item
Bartosiewicz, Z., ``Minimal polynomial realizations,'' {\it Math.\ Control
Signals Systems} {\bf 1}(1988): 227-231.

\item
Basile, G., and G.\ Marro, ``Controlled and conditioned invariant subspaces in
linear system theory,'' {\it J.\ Optimization Theo.\ \& Appl.} {\bf 3}(1969):
306-\-315.

\item
Belevitch, V., {\it Classical Network Theory}, Holden-Day, San Franscisco,
1968.

\item
Bellman, R., {\it Dynamic Programming}, Princeton University Press, Princeton,
1957.

\item
Bellman, R.E., {\it Introduction to the Mathematical Theory of Control
Processes}, Academic Press, New York, 1967.

\item
Bellman, R.E., and R.\ Kalaba (eds.), {\it Selected Papers on Mathematical
Trends in Control Theory}, Dover, New York, 1964.

\item
Bellman, R.E., and R.\ Kalaba, {\it Dynamic Programming and Modern Control
Theory}, Academic Press, New York, 1965.

\item
Bensoussan, A., E.G.\ Hurst, and B.\ Naslund, {\it Management Applications of
Modern Control Theory}, Elsevier, New York, 1974.

\item
Benveniste, A., and P.\ Le Guernic, ``Hybrid dynamical systems theory and the
SIGNAL language,'' {\it IEEE Trans.\ Autom.\ Control} {\bf AC-35}(1990):
535-546.

\item
Berger, M.S., and Berger, M.S., {\it Perspectives in Nonlinearity: An
Introduction to Nonlinear Analysis}, Benjamin, New York, 1968.

\item
Berkovitz, L.D., {\it Optimal Control Theory}, Springer-Verlag, New York,
1974.

\item
Bhatia, N.P., and G.P.\ Szeg\"o, {\it Stability Theory in Dynamical Systems},
Springer-Verlag, Berlin, 1970.

\item
Boltyansky, V.G., ``Sufficient conditions for optimality and the justification
of the dynamic programming method,'' {\it SIAM J.\ Control \& Opt.} {\bf
4}(1966): 326-361.

\item
Bonnard, B., ``Controlabilit\'e des syst\`emes non lin\'eaires,'' {\it C.R.\
Acad.\ Sci.\ Paris, Serie I,} {\bf 292}(1981): 535-537.

\item
Booth, T.L., {\it Sequential Machines and Automata Theory}, Wiley, New York,
1967.

\item
Boothby, W. M., {\it An Introduction to Differentiable Manifolds and
Riemannian Geometry}, Academic Press, New York, 1975.

\item
Boothby, W.M., and R.\ Marino, ``Feedback stabilization of planar nonlinear
systems,'' {\it Systems and Control Letters} {\bf 12}(1989): 87-92.

\item
Boothby, W.M., and E.N.\ Wilson, ``Determination of the transitivity of
bilinear systems,'' {\it SIAM J.Control \& Opt.} {\bf 17}(1979): 212-221.

\item
Bose, N.K., ``Problems and progress in multidimensional systems theory,'' {\it
Proceedings IEEE} {\bf 65}(1976): 824-840.

\item
Bose, N.K., {\it Multidimensional Systems Theory}, Reidel, Dordrecht, 1985.

\item
Boyd, S., V.\ Balakrishnan, and P.\ Kabamba, ``On computing the $H_\infty$
norm of a transfer matrix,'' {\it Math.\ Control Signals Systems} {\bf 2}
(1989): 207-219.

\item
Brammer, R., ``Controllability in linear autonomous systems with positive
controllers,'' {\it SIAM J.\ Control} {\bf 10}(1972): 339-353.

\item
Brewer, J.W., J.W.\ Bunce, and F.S.Van Vleck, {\it Linear Systems over
Commutative Rings}, Marcel Dekker, New York, 1986.

\item
Brewer, J.W., and L.\ Klinger, ``Pole shifting for families of systems: the
dimension one case,'' {\it Math.\ Control Signals Systems} {\bf 1}(1988):
285-292.

\item
Brockett, R.W., ``The status of stability theory for deterministic systems,''
{\it IEEE Trans.\ Autom.\ Control} {\bf AC-11}(1966): 596-606.

\item
Brockett, R.W., ``Structural properties of the equilibrium solutions of
Ricatti equations,'' in {\it Lect.Notes in Math., vol.\ 132} (A.V.\
Balakrishnan et al., eds.), Springer-Verlag, Heidelberg, 1970,pp.\ 61-69.

\item
Brockett, R.W., {\it Finite Dimensional Linear Systems}, Wiley, New York,
1970.

\item
Brockett, R.W., ``On the algebraic structure of bilinear systems,'' in {\it
Theory and Applications of Variable Structure Systems} (R.\ Mohler and A.\
Ruberti, eds.,) Academic Press, New York, 1972,pp.\ 153-168.

\item
Brockett, R.W., ``System theory on group manifolds and coset spaces,'' {\it
SIAM J.\ Control} {\bf 10}(1972): 265-284.

\item
Brockett, R.W., ``Lie theory and control systems defined on spheres,'' {\it
SIAM J.\ Control \& Opt.} {\bf 25}(1973): 213-225.

\item
Brockett, R.W., ``Volterra series and geometric control theory,'' {\it
Automatica} {\bf 12}(1976): 167-176.

\item
Brockett, R.W., ``Nonlinear systems and differential geometry,'' {\it Proc.\
IEEE} {\bf 64}(1976): 61-72.

\item
Brockett, R.W., ``Feedback invariants for nonlinear systems,'' {\it IFAC
Congress}, Helsinki, 1978.

\item
Brockett, R.W., ``Asymptotic stability and feedback stabilization,'' in {\it
Differential Geometric Control theory} (R.W.\ Brockett, R.S.\ Millman, and
H.J. Sussmann, eds.), Birkhauser, Boston, 1983, pp.\ 181-191.

\item
Brockett, R.W., R.S.\ Millman, and H.J.\ Sussmann, eds., {\it Differential
Geometric Control Theory : Proceedings of the Conference Held at Michigan
Technological University, June 28-July 2, 1982}, Birkhauser, Boston, 1983.

\item
Brockett, R.W., and J.L.\ Willems, ``Discretized partial differential
equations: Examples of control systems defined on modules,'' {\it Automatica}
{\bf 10}(1974): 507-515.

\item
Brockett, R.W., and A.S.\ Willsky, ``Finite group homomorphic sequential
systems,'' {\it IEEE Trans.\ Autom.\ Control} {\bf AC-17}(1972): 483-490.

\item
Brogan, W.L., {\it Modern Control Theory}, Prentice-Hall, Englewood Cliffs,
1985.

\item
Brunovsky, P., ``A classification of linear controllable systems,'' {\it
Kybernetika} {\bf 6}(3)(1970): 173-188.

\item
Bumby, R., E.D.\ Sontag, H.J.\ Sussmann, and W.\ Vasconcelos, ``Remarks on the
pole-shifting problem over rings,'' {\it J.\ Pure Appl.\ Algebra} {\bf
20}(1981): 113-127.

\item
Byrnes, C.I., and A.\ Isidori, ``New results and counterexamples in nonlinear
feedback stabilization,'' {\it Systems and Control Letters} {\bf 12}(1989):
437-442.

\item
Byrnes, C.I., ``On the control of certain infinite-dimensional systems by
algebro-geometric techniques,'' {\it Amer.\ J.\ Math.} {\bf 100}(1978):
1333-1381.

\item
Campbell, S.L., {\it Singular Systems of Differential Equations, I} and {\it
II}, Pitman, Marshfield, 1980, 1982.

\item
Campbell, S.L., ``Control problem structure and the numerical solution of
linear singular systems,'' {\it Math.\ Control Signals Systems} {\bf 1}(1988):
73-87.

\item
Canudas de Wit, C.A., {\it Adaptive Control for Partially Known Systems :
Theory and Applications}, Elsevier, New York, 1988.

\item
Casdagli, M., ``Nonlinear prediction of chaotic time series,'' {\it Physica D}
{\bf 35}(1989): 335-356.

\item
Catlin, D.E., {\it Estimation, Control, and the Discrete Kalman Filter},
Springer-Verlag, New York, 1989.

\item
Celle, F., J.P.\ Gauthier, D.\ Kazakos, and G.\ Sallet, ``Synthesis of
nonlinear observers: A harmonic analysis approach,'' {\it Math.\ Systems
Theory (Special Issue on Mathematical Control Theory)} {\bf 22}(1989):
291-322.

\item
Chen, C.T., {\it Introduction to Linear System Theory}, Holt, Reinhart and
Winston, New York, 1970.

\item
Chen, K.T., ``Iterated path integrals,'' {\it Bulletin AMS} {\bf 83}(1977):
831-\-879.

\item
Cho, H., and S.I.\ Marcus, ``On supremal languages of classes of sublanguages
that arise in supervisor synthesis problems with partial observation,'' {\it
Math.\ Control Signals Systems} {\bf 2}(1989): 47-69.

\item
Chow, W.L., ``Uber systeme von linearen partiellen differentialgleichungen
erster ordnung,'' {\it Math.~Ann.} {\bf 117}(1939): 98-105.

\item
Cobb, J.D., ``Feedback and pole placement in descriptor variable systems,''
{\it Intern.\ J.\ Control} {\bf 33}(1981): 1135-1146.

\item
Conte, G., and A.M. Perdon, ``An algebraic notion of zeroes for systems over
rings,'' in {\it Mathematical Theory of Networks and Systems}, (P.A.Fuhrmann,
ed.), Springer-Verlag, Berlin, 1984, pp.\ 323-330.

\item
Constanza, V., B.W. Dickinson, and E.F. Johnson, ``Universal approximations of
discrete-time control systems over finite time,'' {\it IEEE Trans.\ Autom.\
Control} {\bf AC-28}(1983): 439-452.

\item
Conway, J.B., {\it Regular Algebra and Finite Machines}, Chapman and Hall,
London, 1971.

\item
Crandall, M.G., L.C.\ Evans, and P.L.\ Lions, ``Some properties of viscocity
solutions of Ham\-ilton-Jac\-obi equations,'' {\it Trans.\ AMS} {\bf
282}(1984): 487-502.

\item
Craven, B.D., {\it Mathematical Programming and Control Theory}, Wiley, New
York, 1978.

\item
Crouch, P.E., ``Dynamical realizations of finite Volterra series,'' Ph.D.\
Thesis, Harvard, 1977.

\item
Crouch, P.E., ``Spacecraft attitude control and stabilization: applications of
geometric control theory,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 29}(1984):
321-333.

\item
Crouch, P.E., and A.J.\ van der Schaft, {\it Variational and Hamiltonian
Control Systems}, Springer-Verlag, New York, 1987.

\item
Curtain, R., and A.J.\ Pritchard, {\it Infinite Dimensional Linear Systems
Theory}, Springer-Verlag, Berlin, 1978.

\item
Davis, M.H.A., {\it Linear Estimation and Stochastic Control}, Chapman and
Hall, London, 1977.

\item
Dayawansa, W.P., and C.F.\ Martin, ``Asymptotic stabilization of two
dimensional real-analytic systems,'' {\it Systems Control Lett.} {\bf
12}(1989): 205-211.

\item
de Jong, L.S., {\it Numerical Aspects of Realization Algorithms in Linear
Systems Theory}, Ph.D.\ Thesis, Univ.\ of Eindhoven, Mathematics Department,
1975.

\item
Delchamps, D.F., ``Analytic stabilization and the algebraic Ricatti
equation,'' in {\it Proc.\ IEEE Conf.\ Dec.and Control, Dec.\ 1983}, IEEE
Publications, pp.\ 1396-1401.

\item
Delfour, M.C., C.\ McCalla, and S.K.\ Mitter, ``Stability and the
infinite-time quadratic cost problem for linear hereditary differential
systems,'' {\it SIAM J.\ Control} {\bf 13}(1975): 48-88.

\item
Desoer, C.A., {\it Notes for a Second Course on Linear Systems}, Van Nostrand,
New York, 1970.

\item
Desoer, C.A., and M.\ Vidyasagar, {\it Feedback Synthesis: Input-Output
Properties}, Academic Press, New York, 1975.

\item
Desoer, C.A., and M.G.\ Kabuli, ``Nonlinear plants, factorizations, and stable
feedback systems,'' in {\it Proc.\ IEEE Conf.\ Decision and Control}, Los
Angeles, Dec 1987, pp.\ 155-156.

\item
Dickinson, B.W., ``On the fundamental theorem of linear state variable
feedback,'' {\it IEEE Trans.\ Autom.Cntr.}{\bf AC-19}(1974):577-579.

\item
Dickinson, B.W., ``A new characterization of canonical convolutional
encoders,'' {\it IEEE Trans.\ Inform. Theory} {\bf IT-20}(1976): 352-354.

\item
Dickinson, B.W., ``The classification of linear state variable control laws,''
{\it SIAM J.\ Control \& Opt.} {\bf 14}(1976): 467-477.

\item
Dickinson, B.W., and E.D. Sontag, ``Dynamical realizations of sufficient
sequences,'' {\it IEEE Trans.\ Inform.Th.}{\bf IT-31}(1985):670-676.

\item
Dieudonn\'e, J., {\it Foundations of Modern Analysis}, Academic Press, New
York, 1960.

\item
Dolezal, V., {\it Monotone Operators and Applications in Control and Network
Theory}, Elsevier, New York, 1979.

\item
Dorato, P.\ (ed.), {\it Robust Control}, IEEE Press, New York, 1987.

\item
Doyle, J.C., K.\ Glover, P.P.\ Khargonekar, and B.A.\ Francis, ``State-space
solutions to the standard $H_2$ and $H_\infty$ control problems,'' {\it IEEE
Trans.\ Autom.\ Control} {\bf AC-34}(1989): 831-846.

\item
Eilenberg, S., {\it Automata, Languages, and Machines, Volume A}, Academic
Press, New York, 1974.

\item
El-Assoudi, R., and J.P.\ Gauthier, ``Controllability of right-invariant
systems on real simple Lie groups of type $F_4$, $G_2$, $C_n$, and $B_n$,''
{\it Math.\ Control Signals Systems} {\bf 1}(1988): 293-301

\item
Elliott, D.L., ``A consequence of controllability,'' {\it J.\ Diff.\ Eqs.}
{\bf 10} (1971): 364-370.

\item
Emre, E., ``On necessary and sufficient conditions for regulation of linear
systems over rings,'' {\it SIAM J.Control \& Opt.} {\bf 20}(1982): 155-160.

\item
Emre, E., and P.P.\ Khargonekar, ``Regulation of split linear systems over
rings: coefficient-assignment and observers,'' {\it IEEE Trans.\ on Autom.\
Control} {\bf AC-27}(1982): 104-113.

\item
Farmer, J.D., and J.J.\ Sidorowich, ``Exploiting chaos to predict the future
and reduce noise,'' in {\it Evolution, Learning, and Cognition} (Y.C.\ Lee,
ed.)  World Scientific Press, 1988, p. 277ff.

\item
Faurre, P., ``Stochastic realization algorithms,'' in {\it System
Identification - Advances and Case Studies} (R.K.\ Mehra and D.G.\ Lainiotis,
eds.), Academic Press, New York, 1976, pp.\ 1-25.

\item
Faurre, P., M.\ Clerget, and F.\ Germain, {\it Operateurs Rationnels
Positifs}, Dunod, Paris, 1979.

\item
Feintuch, A., and R.\ Saeks, {\it System Theory.  A Hilbert Space Approach},
Academic Press, New York, 1982.

\item
Fleming, W.H., and R.W.\ Rishel, {\it Deterministic and Stochastic Optimal
Control}, Springer-Verlag, New York, 1975.

\item
Fleming, W.H., et. al., {\it Future Directions in Control Theory: A
Mathematical Perspective}, SIAM Publications, Philadelphia, 1988.

\item
Fliess, M., ``Sur certaines familles de s\'eries formelles,'' Th\`ese de
Doctorat d'Etat, Universit\'e Paris VII, 1972.

\item
Fliess, M., ``Matrices de Hankel,'' J.\ Math Pures Appl.{\bf 53}(1974):
197-224.

\item
Fliess, M., ``Sur la realization des systemes dynamiques bilineaires,'' {\it
C.R.\ Acad.\ Sc.\ Paris} {\bf A277}(1973): 243-247.

\item
Fliess, M., ``The unobservability ideal for nonlinear systems,'' {\it IEEE
Trans.\ Autom.\ Control} {\bf AC-26}(1981): 592-593.

\item
Fliess, M., ``Fonctionnelles causales non lin\'eaires et ind\'etermin\'ees non
commutatives,'' {\it Bull.\ Soc.\ Math.\ France} {\bf 109}(1981): 3-40.

\item
Fliess, M., ``Quelques d\'efinitions de la th\'eorie des syst\`emes
\`a la lumi\`ere des corps diff\`erentiels,'' {\it C.R.Acad.\ Sc.\ Paris}
{\bf I}(1987): 91-93.

\item
Fliess, M., and I.\ Kupka, ``A finiteness criterion for nonlinear input-output
differential systems,'' {\it SIAM J.Control \& Opt.} {\bf 21}(1983): 721-728.

\item
Fornasini, E., and G.\ Marchesini, ``Algebraic realization theory of bilinear
discrete-time input/output maps,'' {\it J.Franklin Inst.} {\bf 301}(1976):
143-159.

\item
Fortmann,T.E., and K.L.Hitz, {\it An Introduction to Linear Control Systems},
Marcel Dekker, New York, 1977.

\item
Francis, B.A., {\it A Course in $H_\infty$ Control Theory}, Springer-Verlag,
New York, 1987.

\item
Fruechte, R.D., {\it Application of Control Theory in the Automotive
Industry}, Inderscience Enterprises, Geneve, 1983.

\item
Fuhrmann, P.A., {\it Linear Systems and Operators in Hilbert Space},
McGraw-Hill, New York, 1981.

\item
Fujiwara, H., ``Computational complexity of controllability/obs\-erv\-ability
problems for combinatorial circuits,'' in {\it Proc.\ 18th IEEE Int.\ Symp.\
on Fault Tolerant Computing}, IEEE Publ., 1988, pp.\ 64-69.

\item
Fuller, A.T., ``The early development of control theory, parts I and II,''
{\it J.\ Dynamic Systems, Measurement, and Control} {\bf 98}(1976): 109-118
and 224-235.

\item
Gamkrelidze, R.V., {\it Principles of Optimal Control Theory}, Plenum Press,
New York, 1978.

\item
Gantmacher, F.R., {\it Theory of Matrices}, (English transl.,) Chelsea, New
York, 1959.

\item
Gelb, A., et. al., {\it Optimal Estimation}, MIT Press, Cambridge, MA, 1974.

\item
Gibson, J.A., and T.T.\ Ha, ``Further to the preservation of controllability
under sampling,'' {\it Int.\ J.\ Control} {\bf 31}(1980): 1013-1026.

\item
Gilbert, E., ``Controllability and observability in multivariable control
systems,'' {\it SIAM J.\ Control} {\bf 1}(1963): 128-151.

\item
Ginsburg, S., {\it An Introduction to Mathematical Machine Theory},
Addison-Wesley, Reading, MA, 1962.

\item
Glass, M., ``Exponential stability revisited,'' {\it Int.\ J.\ Control} {\bf
46}(1987): 1505-1510.

\item
Gopal, M., {\it Modern Control System Theory}, Wiley, New York,1984.

\item
Gopinath, B., ``On the control of multiple input-output systems,'' {\it Bell
Systems Tech.\ J.}, March 1971.

\item
Grasse, K.A., ``Perturbations of nonlinear controllable systems,'' {\it SIAM
J.\ Control \& Opt.} {\bf 19}(1981): 203-220.

\item
Grasse, K.A., ``Nonlinear perturbations of control-semilinear control
systems,'' {\it SIAM J.\ Control \& Opt.} {\bf 20}(1982): 311-327.

\item
Grasse, K.A., ``On accessibility and normal accessibility: the openness of
controllability in the fine $C^0$ topology,'' {\it J.Diff.\ Eqs.} {\bf
53}(1984): 387-414.

\item
Grasselli, O.M., and A.\ Isidori, ``An existence theorem for observers of
bilinear systems,'' {\it IEEE Trans.\ Autom.Control} {\bf
26}(1981):1299-\-1301.

\item
Grimble, M.J., and M.A.\ Johnson, {\it Optimal Control and Stochastic
Estimation : Theory and Applications}, Wiley, New York, 1988.

\item
Grizzle, J.W., ``Controlled invariance for discrete-time nonlinear systems
with an application to the decoupling problem,'' {\it IEEE Trans.\ Autom.\
Control} {\bf AC-30}(1985): 868-874.

\item
Grizzle, J.W., and A.\ Isidori, ``Block noninteracting control with stability
via static state feedback,'' {\it Math.\ Control Signals Systems} {\bf
2}(1989): 315-341

\item
Grodins, F.S., {\it Control Theory and Biological Systems.}, Columbia
University Press, New York, 1963.

\item
Hahn, W., {\it Stability of Motion}, Springer-Verlag, New York, 1967.

\item
Hammer, J., ``Fraction representations of nonlinear systems: a simplified
approach,'' {\it Int.\ J.\ Control} {\bf 46}(1987): 455-472.

\item
Hammer, J., ``On nonlinear systems, additive feedback, and rationality,'' {\it
Int.\ J.\ Control} {\bf 40}(1984): 953-969.

\item
Harris, C.J., and Billings, S.A., {\it Self-Tuning and Adaptive Control :
Theory and Applications}, P.\ Peregrinus, London, 1985.

\item
Harvey, C.A., and E.B.\ Lee, ``On stability with linear feedback control,''
Honeywell Corp.\ Memo Number 8150 (to C.R.\ Stone), January 8, 1963.

\item
Hautus, M.L.J., ``Controllability and observability conditions for linear
autonomous systems,'' {\it Ned.\ Akad.\ Wetenschappen, Proc.\ Ser.\ A} {\bf
72}(1969): 443-448.

\item
Hautus, M.L.J., ``Stabilization, controllability, and observability for linear
autonomous systems,'' {\it Ned.\ Akad.\ Wetenschappen, Proc.\ Ser.\ A} {\bf
73}(1970): 448-455.

\item
Hautus, M.L.J., ``A simple proof of Heymann's Lemma,'' Memo COSOR 76-17,
Department of Mathematics, Eindhoven Univ.of Technology, Nov.76.

\item
Hautus, M.L.J., ``$(A,B)$-invariant and stabilizability subspaces, a frequency
domain description,'' {\it Automatica} {\bf 16}(1980): 703-707.

\item
Hautus, M.L.J., and E.D.\ Sontag, ``New results on pole-shifting for
parametrized families of systems'', {\it J.Pure Appl.\ Algebra} {\bf
40}(1986): 229-244.

\item
Hazewinkel, M., and A-M.Perdon, ``On families of systems:
pointwise-local-global isomorphism problems,'' {\it Int.\ J.\ Control} {\bf
33}(1981): 713-\-726.

\item
Helmke, U., ``Linear dynamical systems and instantons in Yang-Mills theory,''
{\it IMA J.Math.\ Control and Information} {\bf 3}(1986): 151-166.

\item
Hermann, R., ``On the accessibility problem in control theory,'' in {\it
Int.~Symp.~Diff.~Eqs.\ and Mechanics}, Academic Press, New York, 1973.

\item
Hermann, R., {\it Geometric Structure Theory of Systems-Control, Theory and
Physics}, Math Sci Press, Brookline, MA, 1974.

\item
Hermann, R., {\it Cartanian Geometry, Nonlinear Waves, and Control Theory},
Math Sci Press, Brookline, MA, 1979.

\item
Hermann, R., and A.J.\ Krener, ``Nonlinear controllability and
observability,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 22}(1977): 728-740.

\item
Hermes, H., ``On a Stabilizing Feedback Attitude Control,'' {\it J.  Optim.\
Theory and Appl.} {\bf 31}(1980): 373-384.

\item
Hermes, H., ``On the synthesis of a stabilizing feedback control via Lie
algebraic methods,'' {\it SIAM J.\ Control \& Opt.} {\bf 18}(1980):352-361.

\item
Hermes, H., and J.P.\ LaSalle, {\it Functional Analysis and Time-Optimal
Control}, Academic Press, New York, 1969.

\item
Hestenes, M.R., {\it Calculus of Variations and Optimal Control Theory}, R.\
E.\ Kriegerz, Huntington, New York, 1980.

\item
Hewitt, E., and K.\ Stromberg, {\it Real and Abstract Analysis},
Springer-Verlag, New York, 1965.

\item
Heymann, M., ``Comments on `Pole assignment in multi-input cont\-rollable
linear systems','' {\it IEEE Trans.\ Autom.\ Control} {\bf AC-13} (1968):
748-749.

\item
Hinrichsen, D., and J.\ O'Halloran, ``A complete characterization of orbit
closures of controllable singular systems under restricted system
equivalence,'' {\it SIAM J. Control \& Opt.} {\bf 28}(1990): 602-623.

\item
Hirschorn, R.M., ``Controllability in nonlinear systems,'' {\it J.\ Diff.\
Eqs.} {\bf 19}(1975): 46-61.

\item
Hirschorn, R.M., ``Invertibility of nonlinear control systems,'' {\it SIAM J.\
Control \& Opt.} {\bf 17}(1979): 289-297.

\item
Hirschorn, R.M., ``(A,B)-invariant distributions and disturbance decoupling in
nonlinear systems,'' {\it SIAM J.\ Control \& Opt.} {\bf 19}(1981): 1-19.

\item
Hunt, L.R., R.\ Su, and G.\ Meyer, ``Design for multi-input nonlinear
systems,'' in {\it Differential Geometric Control theory} (R.W.Brockett,
R.S.Millman, and H.J.Sussmann, eds.), Birkhauser, Boston, 1983.

\item
Isidori, A., ``Direct construction of minimal bilinear realizations from
nonlinear input/output maps,'' {\it IEEE Trans.\ Autom.\ Control} {\bf
18}(1973): 626-631.

\item
Isidori, A., {\it Nonlinear Control Systems: An Introduction},
Springer-Verlag, Berlin, 1985.

\item
Isidori, A., A.J.\ Krener, C.\ Gori-Giorgi, and S.\ Monaco, ``Nonlinear
decoupling via feedback: A differential geometric approach,'' {\it IEEE
Trans.\ Autom.\ Control} {\bf 26}(1981): 331-345.

\item
Jacobson, C.A., and C.N.\ Nett, ``Linear state-space systems in
in\-fin\-ite-dim\-en\-sion\-al space: The role and characterization of joint
stabilizability/detectability,'' {\it IEEE Trans.\ Autom.\ Control} {\bf
33}(1988): 541-549.

\item
Jacobson, D.H., {\it Extensions of Linear-Quadratic Control, Optimization and
Matrix Theory}, Academic Press, New York, 1977.

\item
Jakubczyk, B., ``Invertible realizations of nonlinear discrete time systems,''
in {\it Proc.\ Princeton Conf.Inf.\ Sc.and Systs.}(1980): 235-239.

\item
Jakubczyk, B., ``Poisson structures and relations on vector fields and their
Hamiltonians,'' {\it Bull.\ Polish Acad Sci} {\bf 34}(1986): 713-721.

\item
Jakubczyk, B., ``Feedback linearization of discrete-time systems,'' {\it
Systems and Control Letters} {\bf 9}(1987): 441-416.

\item
Jakubczyk, B., and D.\ Normand-Cyrot, ``Orbites de pseudo-groupes de
diffeomorphismes et commandabilit\'e des systemes non linearires en temps
discret,'' {\it C.R.Acad.\ Sc.\ Paris} {\bf 298-I}(1984): 257-260.

\item
Jakubczyk, B., and W.\ Respondek, ``On linearization of control systems,''
{\it Bull.\ Acad.~Pol.~Sci., Ser.\ Sci.\ Math.\ Astr.\ Phys.} {\bf 28}(1980):
517-522.

\item
Jakubczyk, B., and E.D.\ Sontag, ``Controllability of nonlinear
dis\-crete-time systems: A Lie-algebraic approach,'' {\it SIAM J.\ Control \&
Opt.} {\bf 28}(1990): 1-33.

\item
James, M.R., ``Asymptotic nonlinear filtering and large deviations with
application to observer design,'' Systems Research Center Report TR-88-28,
University of Maryland, 1988.  (Ph.\ D.\ Dissertation.)

\item
Jurdjevic, V., and J.P.\ Quinn, ``Controllability and stability,'' {\it J.\
Diff.Eqs.} {\bf 28}(1978): 381-389.

\item
Jurdjevic, V., and H.J.\ Sussmann, ``Control systems on Lie groups,'' {\it
J.Diff.Eqs.} {\bf 12}(1972): 313-329.

\item
Kailath,T., {\it Linear Systems}, Prentice Hall, Englewood Cliffs, 1980.

\item
Kalman, R.E., ``Nonlinear aspects of sampled-data control systems,'' in {\it
Proc.\ of Symp. on Nonlinear Circuit Analysis} (J. Fox, ed.), Polytechnic
Institute of Brooklyn, 1956, pp.\ 273-313.

\item
Kalman, R.E., ``A new approach to linear filtering and prediction problems,''
{\it ASME Trans.} {\bf 82D}(1960): 33-45.

\item
Kalman, R.E., ``Contributions to the theory of optimal control,'' {\it
Bol.~Soc.~Mat.~Mex.} {\bf 5}(1960): 102-119.

\item
Kalman, R.E., ``On the general theory of control systems,'' in {\it Proc.
First IFAC Congress Automatic Control}, Moscow, 1960, Butterworths, London,
vol. 1, pp.\ 481-492.

\item
Kalman, R.E., ``Canonical structure of linear dynamical systems,'' {\it Proc.\
Natl.\ Acad.\ Sci.} {\bf 48}(1962): 596-600.

\item
Kalman, R.E., ``Mathematical description of linear dynamical systems,'' {\it
SIAM J.\ Control} {\bf 1}(1963): 152-192.

\item
Kalman, R.E., ``The theory of optimal control and the calculus of
variations,'' in {\it Mathematical Optimization Techniques} (R. Bellman, ed.),
Univ. of California Press, 1963, pp.\ 309-331.

\item
Kalman, R.E., ``Algebraic structure of linear dynamical systems, I.  The
module of $\Sigma$,'' {\it Proc. Natl.\ Acad.\ Sci.\ (USA)} {\bf 54}(1965):
1503-1508.

\item
Kalman, R.E., ``Irreducible realizations and the degree of a rational
matrix,'' {\it J.\ SIAM} {\bf 13}(1965): 520-544.

\item
Kalman, R.E., ``Linear stochastic filtering - Reappraisal and outlook,'' in
{\it Proc.\ Symp.\ on System Theory} (J.\ Fox, ed.), Polytechnic Press, New
York, 1965, pp.\ 197-205.

\item
Kalman, R.E., ``Algebraic aspects of the theory of dynamical systems,'' in
{\it Differential Equations and Dynamical Systems} (J.K.\ Hale and J.P.\
LaSalle, eds.), Academic Press, New York, 1967,pp.133-146.

\item
Kalman, R.E., ``Kronecker invariants and feedback,'' in {\it Ordinary
Differential Equations} (L.\ Weiss, ed.), Academic Press, New York, 1972, pp.\
459-471.

\item
Kalman, R.E., ``Algebraic-geometric description of the class of linear systems
of constant dimension,'' in {\it Proc. 8th Princeton Conf.\ Inform.\ Sci.\ and
Systems}, 1974, pp.\ 189-191.

\item
Kalman, R.E., ``Optimization, mathematical theory of, control theory'' in {\it
Encyclopaedia Britannica, Fifteenth Edition}, 1974, pp.\ 643-638.

\item
Kalman, R.E., ``Algebraic aspects of the generalized inverse,'' in {\it
Generalized Inverses and Applications} (M. Zuhair Nashed, ed.), Academic
Press, New York, 1976, pp.\ 111-124.

\item
Kalman, R.E., ``Identifiability and modeling in econometrics,'' {\it
Developments in Statistics} {\bf 4}(1983): 97-135.

\item
Kalman, R.E., and R.S.\ Bucy, ``New results in linear filtering and prediction
theory,'' {\it Trans.\ ASME (J.\ Basic Engineering)} {\bf 83 D}(1961): 95-108.

\item
Kalman, R.E., P.L.\ Falb, and M.A.\ Arbib, {\it Topics in Mathematical System
Theory}, McGraw-Hill, New York, 1969.

\item
Kalman, R.E., Y.C.\ Ho, and K.S.\ Narendra, ``Controllability of linear
dynamical systems,'' {\it Control ~Diff.~Eqs.} {\bf 1}(1963): 189-213.

\item
Kalouptsidis, N., and J.\ Tsinias, ``Stability improvement of nonlinear
systems by feedback,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 29}(1984):
346-367.

\item
Kamen, E.W., ``On an algebraic theory of systems defined by convolution
operators,'' {\it Math.\ Systems Th.} {\bf 9}(1975): 57-74.

\item
Kamen, E.W., ``Module structure of infinite-dimensional systems with
applications to controllability,'' {\it SIAM J.Control \& Opt.} {\bf
14}(1976): 389-408.

\item
Kamen, E.W., ``Lectures on algebraic system theory: linear systems over
rings,'' NASA Contractor Report \#3016, 1978.

\item
Kamen, E.W., and P.K.\ Khargonekar, ``On the control of linear systems
depending on parameters,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 29}(1984):
25-33.

\item
Kamen, E.W., P.K.\ Khargonekar, and K.R.\ Polla, ``A transfer function
approach to linear time-varying systems,'' {\it SIAM J.Control \& Opt.} {\bf
23}(1985): 550-565.

\item
Kappel, F., K.\ Kunisch, and W.\ Schappacher, {\it Control Theory for
Distributed Parameter Systems and Applications}, Springer-Verlag, New York,
1983.

\item
Kawski, M., ``Stabilization of nonlinear systems in the plane,'' {\it Systems
and Control Letters} {\bf 12}(1989): 169-176.

\item
Kenney, C., and A.J.\ Laub, ``Controllability and Stability Radii for
Companion Form Systems,'' {\it Math.\ Control Signals Systems} {\bf 1}(1988):
239-256.

\item
Khargonekar, P.P., ``On matrix fraction representations for linear systems
over commutative rings,'' {\it SIAM J. Control \& Opt.} {\bf 20}(1982):
172-197.

\item
Khargonekar, P.P., T.T.\ Georgiou, and A.M.\ Pascoal, ``On the robust
stabilizability of linear time invariant plants with unstructured
uncertainty,'' {\it IEEE Trans.\ Autom.\ Control} {\bf AC-32}(1987): 201-208.

\item
Khargonekar, P.P., K. R. Poolla, and A. Tannenbaum, ``Robust control of linear
time-invariant plants by periodic compensation,'' {\it IEEE Trans.\ Autom.\
Control} {\bf AC-30}(1985): 1088-1096.

\item
Khargonekar, P.P., and E.D.\ Sontag, ``On the relation between stable matrix
fraction decompositions and regulable realizations of systems over rings,''
{\it IEEE Trans.\ Autom.\ Control} {\bf 27}(1982): 627-638.

\item
Kirk, D.E., {\it Optimal Control Theory; an Introduction}, Prentice-Hall,
Englewood Cliffs, 1970.

\item
Knobloch, H.W., {\it Higher Order Necessary Conditions in Optimal Control
Theory}, Springer-Verlag, New York, 1981.

\item
Koditschek, D.E., ``Adaptive techniques for mechanical systems,'' in {\it
Proc.5th.\ Yale Workshop on Adaptive Systems}, pp.\ 259-265, Yale University,
New Haven, 1987.

\item
Kokotovic, P.V., and H.J.\ Sussmann, ``A positive real condition for global
stabilization of nonlinear systems,'' {\it Systems Control Letters} {\bf
13}(1989): 125-134.

\item
Koppel, L.B., {\it Introduction to Control Theory, with Applications to
Process Control}, Prentice-Hall, Englewood Cliffs, 1968.

\item
Kou, S.R., D.L.\ Elliott, and T.J.\ Tarn, ``Exponential observers for
nonlinear dynamical systems,'' {\it Inform.Contr} {\bf 29}(1975): 204-216.

\item
Krasnoselskii, M.A., and P.P.\ Zabreiko, {\it Geometric Methods of Nonlinear
Analysis}, Springer-Verlag, Berlin, 1983.

\item
Krener, A.J., ``A generalization of Chow's theorem and the bang-bang theorem
to nonlinear control systems,'' {\it SIAM J.\ Control} {\bf 12}(1974): 43-52.

\item
Krener, A.J., ``The high order maximal principle and its application to
singular extremals,'' {\it SIAM J.Control \& and Opt.} {\bf 15}(1977):
256-293.

\item
Krener, A.J., ``(Adf,g), (adf,g) and locally (adf,g) Invariant and
Controllability Distributions,'' {\it SIAM J.Control \& Opt.} {\bf 23}(1985):\
523-549.

\item
Krener, A.J., ``Nonlinear controller design via approximate normal forms,'' in
{\it Proc.\ IMA Conf.on Signal Processing, Minneapolis, June-Aug.\ 1988},
Institute for Mathematics and its Applications, 1989.

\item
Kupka, I., and G.\ Sallet, ``A sufficient condition for the transitivity of
pseudo-groups: Application to system theory,'' {\it J.~Diff.~Eqs.} {\bf
47}(1973): 462-470.

\item
Kwakernaak, H., and R. Sivan, {\it Linear Optimal Control Systems}, Wiley, New
York, 1972.

\item
Lang, S., {\it Analysis II}, Addison-Wesley, Reading, MA, 1969.

\item
Langenhop, C.E., ``On the stabilization of linear systems,'' {\it Proc.\ AMS}
{\bf 15}(1964): 735-742.

\item
Lee, E.B., and L.\ Markus, {\it Foundations of Optimal Control Theory}, Wiley,
New York, 1968.

\item
Lee, K.K., and A.\ Arapostathis, ``On the controllability of piecewise linear
hypersurface systems,'' {\it Systems and Control Letters} {\bf 9}(1987):
89-96.

\item
Lee, K.K., and A.\ Arapostathis, ``Remarks on smooth feedback stabilization of
nonlinear systems,'' {\it Systems and Control Letters} {\bf 10} (1988): 41-44.

\item
Lee, H.G., A.\ Arapostathis, and S.I.\ Marcus, ``On the linearization of
discrete time systems,'' {\it Int.\ J.\ Control} {\bf 45}(1987): 1103-1124.

\item
Leigh, J.R., {\it Functional Analysis and Linear Control Theory}, Academic
Press, New York, 1980.

\item
Lenz, K., J.\ Doyle, and P.\ Khargonekar, ``When is a controller $H_\infty$
optimal?''  {\it Math.\ Control Signals Systems} {\bf 1}(1988): 107-122.

\item
Leontaritis, I.J., and S.A.\ Billings, ``Input-output parametric models for
nonlinear systems,'' Parts I and II, {\it Int.\ J.\ Control} {\bf 41} (1985):
303-344.

\item
Levi, M., and F.M.A.\ Salam, eds., {\it Dynamical Systems Approaches to
Nonlinear Problems in Systems and Circuits}, SIAM Publications, Philadelphia,
1988.

\item
Levis, A.H., et.\ al., ``Challenges to control: A collective view.
Report from the workshop held at the University of Santa Clara on
September 18-19, 1986,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 32}(1987):
275-285.

\item
Lewis, F.L., {\it Optimal Estimation : With an Introduction to Stochastic
Control Theory}, Wiley, New York, 1986.

\item
Lewis, F.L., ``A survey of linear singular systems,'' {\it Circuits, Systs.,
and Signal Proc.} {\bf 5}(1986): 3-36.

\item
Lindquist, A., and G.\ Picci, ``Realization theory for multivariate stationary
Gaussian processes,'' {\it SIAM J.Control \& Opt.} {\bf 23}(1985): 809-857.

\item
Lions, J.L., {\it Optimal Control of Systems Governed by Partial Differential
Equations}, Springer-Verlag, Berlin, 1971.

\item
Ljung, L., and T.\ S\"oderstr\"om, {\it Theory and Practice of Recursive
Estimation}, MIT Press, Cambridge, 1983.

\item
Lo, J.T., ``Global bilinearization of systems with controls appearing
linearly,'' {\it SIAM J.Control} {\bf 13}(1975): 879-885.

\item
Lobry, C., ``Controllabilit\'e des syst\`emes non lin\'eaires,'' {\it SIAM J.\
Control} {\bf 8}(1970): 573-605.

\item
Logemann, H., ``On the transfer matrix of a neutral system: Characterizations
of exponential stability in input-output terms,'' {\it Systems and Control
Letters} {\bf 9}(1987): 393-400.

\item
Longchamp, R., ``State-feedback control of bilinear systems,'' {\it IEEE
Trans.\ Autom.\ Cntr.} {\bf 25}(1980): 302-306.

\item
Luenberger, D.G., ``An introduction to observers,'' {\it IEEE Trans.\ Autom.\
Cntr.} {\bf 16}(1971): 596-602.

\item
MacFarlane, A.G.J., ``An eigenvector solution of the optimal linear
regulator,'' {\it J.Electron.\ Control} {\bf 14}(1963): 643-654.

\item
Macki, J., and A.\ Strauss., {\it Introduction to Optimal Control Theory},
Springer-Verlag, New York, 1982.

\item
Manes, E.G., ed., {\it Category Theory Applied to Computation and Control
Proceedings of the First International Symposium, San Francisco, February
25-26, 1974}, Springer-Verlag, New York, 1975.

\item
Manitius, A., ``Necessary and sufficient conditions of approximate
controllability for general linear retarded systems,'' {\it SIAM J.  Control
\& Opt.\/} {\bf 19}(1981): 516-532.

\item
Marcus, S.I., ``Optimal nonlinear estimation for a class of discrete time
stochastic systems,'' {\it IEEE Trans.\ Autom.\ Cntr.} {\bf 24}(1979):
297-302.

\item
Marcus, S.I., ``Algebraic and geometric methods in nonlinear filtering,'' {\it
SIAM J.\ Control \& Opt.} {\bf 22}(1984): 817-844.

\item
Marino, R., ``Feedback stabilization of single-input nonlinear systems,'' {\it
Systems and Control Letters} {\bf 10}(1988): 201-206.

\item
Markus, L., and H.\ Yamabe, ``Global stability criteria for differential
systems,'' {\it Osaka J.\ Math.} {\bf 12}(1960): 305-317.

\item
Massera, J.L., ``Contributions to stability theory,'' {\it Annals of Math}
{\bf 64}(1956): 182-206.  Erratum in {\it Annals of Math} {\bf 68}(1958): 202.

\item
Massey, J.L., and M.K.\ Sain, ``Codes, automata and continuous systems:
Explicit interconnections,'' {\it IEEE Trans.\ Autom.\ Control} {\bf
12}(1970): 644-650.

\item
Mayeda, H., and T.\ Yamada, ``Strong structural controllability,'' {\it SIAM
J.\ Control \& Opt.} {\bf 17}(1979): 123-138.

\item
Mayr, O., {\it The Origins of Feedback Control}, MIT Press, Cambridge, 1970.

\item
Mesarovic, M.D., and Y.Takahara, {\it General Systems Theory: Mathematical
Foundations}, Academic Press, New York, 1975.

\item
Minorsky, N., {\it Theory of Nonlinear Control Systems}, McGraw-Hill, New
York, 1969.

\item
Monaco, S., and D.\ Normand-Cyrot, ``Invariant distributions for nonlinear
discrete-time systems,'' {\it Systems and Control Letters} {\bf 5}(1984):
191-196.

\item
Moog, C.\ H., ``Nonlinear decoupling and structure at infinity,'' {\it Math.\
Control Signals Systems} {\bf 1}(1988): 257-268

\item
Moroney, P., {\it Issues in the Implementation of Digital Feedback
Compensators}, MIT Press, Cambridge, MA, 1983.

\item
Morse, A.S., ``Ring models for delay differential systems,'' {\it Automatica}
{\bf 12}(1976): 529-531.

\item
Morse, A.S., and L.M.\ Silverman, ``Structure of index-invariant systems,''
{\it SIAM J.\ Control} {\bf 11}(1972): 215-225.

\item
Narendra, K.S., and K. Parthasarathy, ``Identification and control of
dynamical systems using neural networks,'' {\it IEEE Trans.\ Neural Nets} {\bf
1}(1990): 4-27.

\item
Nieuwenhuis, J.W., and J.C. Willems, ``Continuity of dynamical systems - a
system theoretic approach,'' {\it Math.\ Control Signals Systems} {\bf
1}(1988): 147-165.

\item
Nijmeijer, H., ``Observability of autonomous discrete-time nonlinear systems:
a geometric approach,'' {\it Int.\ J.\ Control} {\bf 36}(1982): 867-874.

\item
Nijmeijer, H., and A.V.\ Van der Schaft, {\it Nonlinear Dynamical Control
Systems}, Springer-Verlag, New York, 1990.

\item
Ocone, O., ``Probability densities for conditional statistics in the cubic
sensor problem,'' {\it Math.\ Control Signals Systems} {\bf 1}(1988): 183-202.

\item
Ozveren, C.M., G.C.\ Verghese, and A.S.\ Willsky, ``Asymptotic orders of
reachability in perturbed linear systems,'' {\it IEEE Trans.\ Autom.\ Control}
{\bf 33}(1988): 915-923.

\item
Padulo, L., and M.A.\ Arbib, {\it System Theory.  A Unified State-Space
Approach to Continuous and Discrete Systems}, Saunders, Philadelphia, 1974.
Republished by Hemisphere, Washington, DC.

\item
Pallu de La Barriere, R., {\it Optimal Control Theory; A Course in Automatic
Control Theory}, Saunders, Philadelphia, 1967.

\item
Pandolfi, L., ``Canonical realizations of systems with delays,'' {\it SIAM J.\
Control \& Opt.} {\bf 21}(1983): 598-613.

\item
Pitchford, J.D., and S.J.\ Turnovsky, {\it Applications of Control Theory to
Economic Analysis}, Elsevier/North-Holland, Amsterdam, 1977.

\item
Polderman, J.W., ``A state space approach to the problem of adaptive pole
placement,'' {\it Math.\ Control Signals Systems} {\bf 2}(1989): 71-94

\item
Pontryagin, L.S., V.G.\ Boltyanskii, R.V.\ Gamkrelidze, and E.F.\ Mischenko,
{\it The Mathematical Theory of Optimal Processes}, Wiley, New York, 1962.

\item
Popov, V.M., ``On a new problem of stability for control systems,'' {\it
Autom.\ Remote Control} (English translation) {\bf 24}(1963): 1-23.

\item
Popov, V.M., {\it Hyperstabilityy of Control Systems}, Springer-Verlag,
Berlin, 1973. (Translation of {\it Hiperstabilitatea Sistemelor Automate},
published in 1966 in Rumanian.)

\item
Potter, J.E., ``Matrix quadratic solutions,'' {\it SIAM J.\ Appl.  Math.} {\bf
14}(1966): 496-501.

\item
Prichard, A.J., and J.\ Zabczyk, ``Stability and stabilizability of infinite
dimensional systems,'' {\it SIAM Review} {\bf 23}(1983): 25-52.

\item
Ramirez, W.F., {\it Application of Optimal Control Theory to Enhanced Oil
Recovery}, Elsevier, New York, 1987.

\item
Rantala, J., {\it An Application of Stochastic Control Theory to Insurance
Business}, University of Tampere, Tampere, 1984.

\item
Riggs, D.S., {\it Control Theory and Physiological Feedback Mechanisms},
Krieger, Huntington, New York, 1976.

\item
Rissanen, J., ``Control system synthesis by analogue computer based on the
`generalized linear feedback' concept,'' in {\it Proc.\ Symp.\ Analoge Comp.\
Applied to the Study of Chem.\ Processes}, Presses Acad\'emiques
Europ\'eennes, Brussels, 1961.

\item
Rosenbrock, H.H., {\it State Space and Multivariable Theory}, Nelson, London,
1970.

\item
Rouchaleau, Y., B.F.\ Wyman, and R.E.\ Kalman, ``Algebraic structures of
linear dynamical systems III.  Realization theory over a commutative ring,''
{\it Proc.\ Nat.\ Acad.\ Sci.} {\bf 69}(1972): 3404-3406.

\item
Rouchaleau, Y., and B.F.\ Wyman, ``Linear dynamical systems over integral
domains,'' {\it J.\ Comput.\ System Sci.} {\bf 9}(1974): 129-142.

\item
Rouche, N., P.\ Habets, and M.\ Laloy, {\it Stability Theory by Lyapunov's
Direct Method}, Springer-Verlag, New York, 1977.

\item
Rugh, W.J., {\it Mathematical Description of Linear Systems}, Marcel Dekker,
New York, 1975.

\item
Rugh, W.J., {\it Nonlinear System Theory: the Volterra/Wiener Approach}, The
Johns Hopkins University Press, Baltimore, 1981.

\item
Russell, D.L., {\it Mathematics of Finite Dimensional Control Systems}, Marcel
Dekker, New York, 1979.

\item
Ryan, E.P., and N.J.\ Buckingham, ``On asymptotically stabilizing feedback
control of bilinear systems,'' {\it IEEE Trans.\ Autom.\ Cntr.} {\bf
28}(1983): 863-864.

\item
Sage, A.P., and J.L.\ Melsa, {\it Estimation Theory with Applications to
Communications and Control}, Krieger, Huntington, 1979.

\item
Salam, F.M.A., ``Feedback stabilization of the nonlinear pendulum under
uncertainty: a robustness issue,'' {\it Systems and Control Letters} {\bf
7}(1986): 199-206.

\item
Saskena, V.R., J.O'Relly, and P.V.\ Kokotovic, ``Singular perturbations and
two-scale methods in control theory: survey 1976-1983,'' {\it Automatica} {\bf
20}(1984): 273-293.

\item
Schmitendorf, W.E., and B.R.\ Barmish, ``Null controllability of linear
systems with constrained controls,'' {\it SIAM J.\ Control \& Opt.} {\bf
18}(1980): 327-345.

\item
Schwartz, C.A., B.W.\ Dickinson, and E.D.\ Sontag, ``Characterizing innovation
representations for discrete-time random processes,'' {\it Stochastics} {\bf
11}(1984): 159-172.

\item
Seidman, T.I., ``How violent are fast controls?,'' {\it Math.\ Control Signals
Systems} {\bf 1}(1988): 89-95.

\item
Seierstad, A., and K.\ Sydsater, {\it Optimal Control Theory with Economic
Applications}, North-Holland, New York, 1987.

\item
Sethi, S.P., and G.L.\ Thompson, {\it Optimal Control Theory : Applications to
Management Science}, Kluwer, Boston, 1981.

\item
Sharma, P.K., ``On pole assignment problem in polynomial rings,'' {\it Systems
and Control Letters} {\bf 5}(1984): 49-54.

\item
Siljak, D.D., {\it Large Scale Dynamic Systems}, North-Holland, New York,
1978.

\item
Silva Leite, F., and P.E.\ Crouch, ``Controllability on classical Lie
groups,'' {\it Math.\ Control Signals Systems} {\bf 1}(1988): 31-42.

\item
Singh, M.G.\ (editor-in-chief), {\it Systems \& Control Encyclopedia : Theory,
Technology, Applications}, Pergamon Press, Oxford, 1987.

\item
Sira-Ramirez, H., ``A geometric approach to pulse-width-modulated control in
dynamical systems,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 34} (1989):
184-187.

\item
Skowronski, J.M., {\it Control Theory of Robotic Systems}, World Scientific,
Teaneck, 1989.

\item
Slemrod, M., ``Stabilization of bilinear control systems with applications to
nonconservative problems in elasticity,'' {\it SIAM J.Control \& Opt.} {\bf
16}(1978): 131-141.

\item
Sontag, E.D., ``On certain questions of rationality and decidability,'' {\it
J.\ Comp.\ Syst.\ Sci.} {\bf 11}(1975): 375-381.

\item
Sontag, E.D., ``Linear systems over commutative rings: A survey,'' {\it
Ricerche di Automatica} {\bf 7}(1976): 1-34.

\item
Sontag, E.D., ``On the observability of polynomial systems,'' {\it SIAM
J.Control \& Opt.} {\bf 17}(1979):139-151.

\item
Sontag, E.D., ``Realization theory of discrete-time nonlinear systems: Part I-
The bounded case,'' {\it IEEE Trans.\ Circuits and Syst.} {\bf CAS-26}(1979):
342-356.

\item
Sontag, E.D., {\it Polynomial Response Maps}, Springer-Verlag, New York, 1979.

\item
Sontag, E.D., ``Conditions for abstract nonlinear regulation,'' {\it
Information and Control} {\bf 51}(1981): 105-127.

\item
Sontag, E.D., ``Nonlinear regulation: The piecewise linear approach,'' {\it
IEEE Trans.\ Autom.\ Control} {\bf AC-26}(1981): 346-358.

\item
Sontag, E.D., ``A Lyapunov-like characterization of asymptotic
controllability,'' {\it SIAM J.\ Control \& Opt.} {\bf 21}(1983): 462-471.

\item
Sontag, E.D., ``An algebraic approach to bounded controllability of nonlinear
systems,'' {\it Int.\ J.\ Control} {\bf 39}(1984): 181-188.

\item
Sontag, E.D., ``A concept of local observability,'' {\it Systems and Control
Letters} {\bf 5}(1984): 41-47.

\item
Sontag, E.D., ``Controllability is harder to decide than accessibility,'' {\it
SIAM J.Control \& Opt.} {\bf 26}(1988): 1106-1118.

\item
Sontag, E.D., ``Finite dimensional open-loop control generators for nonlinear
systems,'' {\it Int.\ J.\ Control} {\bf 47} (1988): 537-556.

\item
Sontag, E.D., ``Remarks on stabilization and input-to-state stability,'' {\it
Proc.\ IEEE Conf.\ Dec.and Control, Tampa, Dec.\ 1989}, IEEE Publications,
pp.\ 1376-1378.

\item
Sontag, E.D., ``Smooth stabilization implies coprime factorization,'' {\it
IEEE Trans.\ Autom.\ Control} {\bf 34}(1989): 435-443.

\item
Sontag, E.D.,``Integrability of certain distributions associated with actions
on manifolds and applications to control problems,'' in {\it Nonlinear
Controllability and Optimal Control} (H.J. Sussmann, ed.), pp.\ 81-131, Marcel
Dekker, New York 1990.

\item
Sontag, E.D., ``Feedback stabilization of nonlinear systems,'' in {\it Robust
Control of Linear Systems and Nonlinear Control} (M.A. Kaashoek, J.H.\ van
Schuppen, and A.C.M.\ Ran, eds,.)  Birkh\"auser, Cambridge, MA, 1990.

\item
Stengel, R.F., {\it Stochastic Optimal Control : Theory and Application},
Wiley, New York, 1986.

\item
Stephanopoulos, G., {\it Chemical Process Control : An Introduction to Theory
and Practice}, Prentice-Hall, Englewood Cliffs, 1984.

\item
Sussmann, H.J., ``Semigroup representations, bilinear approximations of
input-output maps, and generalized inputs,'' in {\it Mathematical Systems
Theory, Udine 1975} (G.\ Marchesini, Ed.,) Springer-Verlag, New York, pp.\
172-192.

\item
Sussmann, H.J., ``Minimal realizations and canonical forms for bilinear
systems,'' {\it J.\ Franklin Institute} {\bf 301}(1976): 593-604.

\item
Sussmann, H.J., ``Existence and uniqueness of minimal realizations of
nonlinear systems,'' {\it Math.\ Sys.\ Theory}{\bf 10} (1977): 263-284.

\item
Sussmann, H.J., ``Single-input observability of continuous-time systems,''
{\it Math.\ Systems Theory} {\bf 12}(1979): 371-393.

\item
Sussmann, H.J., ``Subanalytic sets and feedback control,'' {\it J.Diff.\ Eqs.}
{\bf 31}(1979): 31-52.

\item
Sussmann, H.J., ``Analytic stratifications and control theory,'' in {\it
Proc.\ 1978 Internat.\ Congr.\ Mathematicians, Helsinki}, pp.\ 865-\-871.

\item
Sussmann, H.J., ``Lie brackets, real analiticity, and geometric control,'' in
{\it Differential Geometric Control theory} (R.W.\ Brockett, R.S.\ Millman,
and H.J.\ Sussmann, eds.), Birkhauser, Boston, 1983.

\item
Sussmann, H.J., ``A general theorem on local controllability,'' {\it SIAM J.\
Control \& Opt.} {\bf 25}(1987): 158-194.

\item
Sussmann, H.J., ``Minimal realizations of nonlinear systems,'' in {\it
Geometric Methods in Systems Theory} (D.Q.\ Mayne and R.W.\ Brockett, eds.),
Reidel, Dordrecht, 1973, pp.\ 243-252.

\item
Sussmann, H.J., and V.\ Jurdjevic, ``Controllability of nonlinear systems,''
{\it J.\ Diff.\ Eqs.} {\bf 12}(1972): 95-116.

\item
Swan, G.W., {\it Applications of Optimal Control Theory in Biomedicine},
Marcel Dekker, New York, 1984.

\item
Tannenbaum, A., {\it Invariance and System Theory: Algebraic and Geometric
Aspects}, Springer-Verlag, New York, 1980.

\item
Tannenbaum, A., ``On pole assignability over polynomial rings,'' {\it Systems
and Control Letters} {\bf 2}(1982):13-16.

\item
Tannenbaum,A., ``Polynomial rings over arbitrary fields in two or more
variables are not pole assignable,'' {\it Systems and Control Letters} {\bf
2}(1982): 222-224.

\item
Tay, T.T., and J.B.\ Moore, ``Left coprime factorizations and a class of
stabilizing controllers for nonlinear systems,'' in {\it Proc.\ IEEE Conf.
Decision and Control}, Austin, Dec.1988, pp.\ 449-454.

\item
Taylor, A.E., {\it Introduction to Functional Analysis}, Wiley, New York,
1958.

\item
Thompson, W.\ (Lord Kelvin), ``Mechanical integration of the general linear
differential equation of any order with variable coefficients,'' {\it Proc.\
Roc.\ Soc.} {\bf 24}(1876): 271-275.

\item
Toates, F.M., {\it Control Theory in Biology and Experimental Psychology},
Hutchinson, London, 1975.

\item
Tsinias, J., ``Sufficient Lyapunovlike conditions for stabilization,'' {\it
Math.\ Control Signals Systems} {\bf 2}(1989): 343-357.

\item
Tsinias, J., and N.\ Kalouptsidis, ``Prolongations and stability analysis via
Lyapunov functions of dynamical polysystems,'' {\it Math.\ Systems Theory}
{\bf 20}(1987): 215-233.

\item
Tsitsiklis, J.N., ``On the control of discrete-event dynamical systems,'' {\it
Math.\ Control Signals Systems} {\bf 2}(1989): 95-107.

\item
Utkin, V.I., {\it Sliding Modes and Their Application in Variable Structure
Systems}, Mir, Moscow, 1978.

\item
Van der Schaft, A.J., ``Stabilization of Hamiltonian Systems,'' {\it J.\
Nonlinear Analysis, MTA} {\bf 10}(1986): 1021-1035.

\item
Van der Schaft, A.J., ``On realizations of nonlinear systems described by
higher-order differential equations,'' {\it Math Systems Theory} {\bf 19}
(1987): 239-275.

\item
Van Schuppen, J.H., ``Stochastic realization problems,'' in {\it Three Decades
of Mathematical System Theory} (Nijmeijer, H., and J.M.\ Schumacher, eds.),
Springer-Verlag, Berlin, 1989.

\item
Varaiya, P.P., and R.\ Liu, ``Bounded-input bounded-output stability of
nonlinear time-varying differential systems,'' {\it SIAM J.Control} {\bf
4}(1966): 698-704.

\item
Veliov, V.M., and M.I.\ Krastanov, ``Controllability of piecewise linear
systems,'' {\it Systems and Control Letters} {\bf 7}(1986): 335-341.

\item
Verma, M.S., ``Coprime fractional representations and stability of nonlinear
feedback systems,'' {\it Int.\ J.\ Control} {\bf 48}(1988): 897-918.

\item
Vidyasagar, M., ``Input-output stability of a broad class of linear
time-invariant multivariable systems,'' {\it SIAM J.Ctr.}{\bf
10}(1972):203-209.

\item
Vidyasagar, M., {\it Nonlinear Systems Analysis}, Prentice-Hall, Englewood
Cliffs, 1978.

\item
Vidyasagar, M., ``Decomposition techniques for large-scale systems with
nonadditive interactions: stability and stabilizability,'' {\it IEEE Trans.\
Autom.\ Control} {\bf 25}(1980): 773-779.

\item
Vidyasagar, M., ``On the stabilization of nonlinear systems using state
detection,'' {\it IEEE Trans.\ Autom.\ Control} {\bf 25}(1980): 504-509.

\item
Vidyasagar, M., {\it Control Systems Synthesis: A factorization Approach}, MIT
Press, Cambridge, MA, 1985.

\item
Vinter, R.B., ``New results on the relationship between dynamic programming
and the maximum principle,'' {\it Math.\ Control Signals Systems} {\bf
1}(1988): 97-105.

\item
Viswanadham, N., and Vidyasagar, M., ``Stabilization of linear and nonlinear
dynamical systems using an observer-controller configuration,'' {\it Systems
and Control Letters} {\bf 1} (1981): 87-91.

\item
Vukobratovic, M., and D.\ Stokic, {\it Control of Manipulation Robots : Theory
and Application}, Springer-Verlag, Berlin, 1982.

\item
Wang, Y., and E.D.\ Sontag, ``Input/output equations and realizability,'' in
{\it Realization and Modelling in System Theory} (M.A.\ Kaashoek, J.H.\ van
Schuppen, and A.C.M.\ Ran, eds.), Birkh\"auser, Cambridge, MA, 1990.

\item
Willems, J.C., {\it The Analysis of Feedback Systems}, MIT Press, Cambridge,
MA, 1971.

\item
Willems, J.C., ``Least squares stationary optimal control and the algebraic
Riccati equation,'' {\it IEEE Trans.\ Autom.\ Control} {\bf AC-16}(1971):
621-634.

\item
Willems, J.C., ``From time series to linear systems, Part I: Finite
dimensional time-invariant systems,''{\it Automatica} {\bf
22}(1986):561-\-580.

\item
Willems, J.C., ``From time series to linear systems, Part II: Exact
Modelling,'' {\it Automatica} {\bf 22}(1986): 675-694.

\item
Willems, J.C., ``From time series to linear systems, Part III: Approximate
Modelling,'' {\it Automatica} {\bf 23}(1987): 87-115.

\item
Willsky, A.S., {\it Digital Signal Processing and Control and Estimation
Theory : Points of Tangency, Areas of Intersection, and Parallel Directions},
MIT Press, Cambridge, MA, 1979.

\item
Wilson, F.Wesley, ``Smoothing derivatives of functions and applications,''
{\it Trans.\ AMS} {\bf 139}(1969): 413-428.

\item
Windeknecht, T.G., {\it General Dynamical Processes}, Academic Press, New
York, 1971.

\item
Wonham, M., {\it Linear Multivariable Control}, Springer-Verlag, New York,
1974.

\item
Wonham, W.M., ``On pole assignment in multi-input controllable linear
systems,'' {\it IEEE Trans.\ Autom.\ Control} {\bf AC-12}(1967):660-665.

\item
Wonham, W.M., and A.S.\ Morse, ``Decoupling and pole assignment in linear
multivariable systems: a geometric approach,'' {\it SIAM J.\ Control} {\bf
8}(1970): 1-18.

\item
Wonham, W.M., and P.J.\ Ramadge, ``Modular supervisory control of discrete
event systems,'' {\it Math.\ Control Signals Systems} {\bf 1}(1988): 13-30.

\item
Wyman, B.F., ``Time-varying linear discrete-time systems - realization
theory,'' in {\it Advances in Math., Supp.\ Studies, Vol. I}, Academic Press,
New York, 1978, 233-258.

\item
Wyman, B.F., ``Time-varying linear discrete-time systems: duality,'' {\it
Pacific J.\ of Math.} {\bf 86}(1980): 361-377.

\item
Wyman, B.F., and M.K.\ Sain, ``The zero module and essential inverse
systems,'' {\it IEEE Trans.\ Circuits and Systems} {\bf CAS-28}(1981):
112-126.

\item
Yamamoto, Y., ``Realization theory of infinite dimensional linear systems,
Parts I,II,'' {\it Math Sys.\ Theory} {\bf 15}(1981): 55-77, 169-190.

\item
Yamamoto, Y., ``Pseudo-rational input/output maps and their realizations: a
fractional representation approach to infinite-dimensional systems,'' {\it
SIAM J.\ Control \& Opt.} {\bf 26-6}(1988): 1415-1430.

\item
Yamamoto, Y., ``Correspondence of internal and external stability -
Realization, transfer functions and complex analysis,'' in {\it Realization
and Modelling in System Theory} (M.A. Kaashoek, J.H.\ van Schuppen, and
A.C.M.\ Ran, eds,.)  Birkh\"auser, Cambridge, MA, 1990.

\item
Yamamoto, Y., and E.D.\ Sontag, ``On the existence of approximately coprime
factorizations for retarded systems,'' {\it Systems and Control Letters} {\bf
13}(1989): 53-58.

\item
Youla, D.C., H.A.\ Jabr, and J.J.\ Bongiorno, ``Modern Wiener-Hopf design of
optimal controllers: part II,'' {\it IEEE Trans.\ Autom.\ Cntr.} {\bf
AC-21}(1976): 319-338.

\item
Young, L.C., {\it Lectures on the Calculus of Variations and Optimal Control
Theory}, Chelsea, New York, 1980.

\item
Zabczyk J., ``Some comments on stabilizability,'' {\it Applied Math Optimiz.}
{\bf 19}(1989): 1-9.

\item
Zadeh, L.A., and C.A.\ Desoer, {\it Linear System Theory --- The State-Space
Approach}, McGraw-Hill, New York, 1963.

\end{enumerate}
\end{document}
